\section{Introduction}\label{intro}

Often, past information carried on by time series data can help one understand future phenomena. Time series forecasting is hereafter attractive in many research fields and lots of research work are developed to improve the prediction ability of the models. Traditionally, autoregressive models and moving average models are often applied to time series prediction. However, they essentially only capture linear relationship in the data set and can not afford many real cases, such as weather forecasting, financial market prediction or signal processing, which are often complex, nonlinear or even chaotic. Another class of statistical methods is Bayesian inference which is leveraged to deal with uncertainty in noisy environment and improve model accuracy by collecting some prior information. For example, Sahoo and Patra \cite{Sahoo18092020} combine the Bayesian inference method with Markov Chain Monte Carlo (MCMC) algorithm to evaluate water pollution rates. With the help of some prior knowledge, this method can greatly improve the prediction accuracy. However, if the prior distribution is not presumed correctly, the inference will go to the wrong direction. Hence, improving the reliability of prior knowledge is crucial to the model performance. As there always exists various kinds of uncertainties in the forecasting problems, how to accurately capture the intrinsic volatile nature of time series data is still challenging.

Financial time series forecasting is particularly difficult because of the irregular and highly dynamic structure of market behaviour. Stock market indices are influenced by a large number of external factors including economic policies, geopolitical events, investor sentiment and unexpected global incidents. Such influences often introduce sudden jumps, heavy tails and irregular fluctuations in the observed data. Consequently, financial markets exhibit strong nonlinearity and stochasticity, making classical deterministic approaches insufficient for modeling their underlying dynamics. Moreover, financial time series are usually non-stationary in nature, meaning that their statistical properties change over time. This characteristic further complicates the task of building robust predictive models capable of generalizing well under changing market conditions.

Traditional econometric models such as ARIMA and GARCH have achieved considerable success in modeling volatility and short term dependencies in financial data. Nevertheless, these approaches generally rely on restrictive assumptions regarding stationarity, Gaussian noise or linear dependence structures. In practical situations, however, market data often demonstrate abrupt regime changes, volatility clustering and asymmetric behaviour that cannot be adequately represented under these assumptions. Therefore, there has been increasing interest in developing more flexible learning based frameworks which can better adapt to the complex dynamics of real financial systems.

Recently, machine learning models outperform conventional methods in many aspects and have led a growing amount of researches to utilize various deep learning frameworks in the research of time series forecasting. Deep neural networks possess strong representation capability and can automatically extract complex nonlinear patterns from high dimensional data. Recurrent neural networks (RNNs), long short term memory networks (LSTMs) and gated recurrent units (GRUs) have been widely applied to sequential modeling tasks due to their ability to preserve temporal information. Convolutional neural networks have also been explored in time series forecasting by treating temporal dependencies as local patterns. Although these methods have demonstrated promising predictive performance, they still face difficulties in explicitly modeling uncertainty and stochastic evolution in dynamical systems.

Nowadays, the partnership between deep learning and dynamical systems has become a topic of increasingly popularity, as they could benefiting each other in both theoretical and computational aspects. On the computational side, for example, Neural ODE \cite{chen2019neuralordinarydifferentialequations} is a combination of residual network and ordinary differential equations. Instead of considering neural networks as discrete layer transformations, Neural ODE models treat the hidden states as continuous trajectories evolving over time. This viewpoint provides a mathematically elegant framework for understanding deep learning dynamics and allows adaptive computation through differential equation solvers. Such continuous time formulations are especially suitable for time series applications where observations are inherently sequential and temporal.

For stochastic cases, neural stochastic differential equation (Neural SDE) \cite{liu2019neuralsdestabilizingneural} shows a continuous neural network framework based on stochastic differential equation with Brownian motions. The core of the SDE-Net is to treat the deep neural network transformation as the state evolution of a stochastic dynamic system, and introduce the Brownian motion term to capture cognitive uncertainty. By incorporating stochasticity directly into the hidden state evolution, Neural SDEs are able to model random fluctuations more naturally than deterministic neural networks. In addition, the Brownian diffusion term provides robustness against noise and improves the generalization ability of the model under uncertain environments. Such characteristics make Neural SDEs particularly attractive in financial forecasting problems where uncertainty is unavoidable.

In many real world applications, Lévy noise plays an important part as a more general random fluctuation including complex non-Gaussian properties. Zhu \cite{ZHU2014126} compares the Lévy noise with the standard Gaussian noise and concludes that the Lévy noise is more versatile and has a wider range of applications in diverse areas. Unlike Brownian motion, Lévy processes are capable of describing discontinuous jumps and impulsive behaviours. The properties of Lévy motion enables it to better capture the uncertainty in financial data, such as abrupt transitions. In financial markets, sudden crashes or rapid upward movements are common phenomena which cannot be accurately characterized solely through Gaussian diffusion processes. Therefore, introducing Lévy driven stochastic dynamics into neural differential equation models may provide a more realistic representation of market behaviour.

Motivated by these observations, this project investigates stochastic differential equation based neural network architectures for financial time series forecasting. In particular, both Brownian motion driven SDE-Net and Lévy motion driven differential equation networks are explored and compared for forecasting stock market indices. The models are trained and evaluated on financial data from NIFTY50 and SSE50 indices under multiple prediction horizons. Different forms of stochastic perturbations are incorporated into the hidden dynamics in order to analyze their impact on prediction accuracy and model stability. Through this study, the effectiveness of Lévy induced dynamics in capturing complex market uncertainty is examined and compared against the traditional Brownian motion framework.

Another important motivation behind this work is to study whether non-Gaussian stochastic processes can improve the capability of deep learning models in learning extreme market behaviours. Since financial data often contain heavy tailed distributions and sudden jumps, it is reasonable to hypothesize that Lévy driven models may achieve better adaptability under such environments. Furthermore, understanding the influence of different stochastic mechanisms may contribute towards the development of more robust neural forecasting frameworks for real world financial applications.

The rest of this report is organized as follows. In Section 2, neural network models based on stochastic differential networks induced by Brownian motion and Lévy motions are presented in detail. The mathematical background of stochastic differential equations and the corresponding neural architectures are also discussed. In Section 3, experimental results with real financial data are reported, including comparative studies under different forecasting horizons and evaluation metrics. Finally, the report concludes in Section 4 with a discussion of the major findings, limitations of the current work and possible directions for future research.

\clearpage