\chapter{Problem Statement}

Financial markets are highly volatile and uncertain systems influenced by numerous economic, political and social factors. Due to the presence of irregular fluctuations, abrupt jumps and nonlinear dependencies, accurately forecasting stock market indices remains a difficult challenge in time series analysis. Traditional statistical approaches such as autoregressive and moving average models generally rely on assumptions of linearity and Gaussian noise, which limit their effectiveness when dealing with real world financial data exhibiting heavy tailed distributions and sudden transitions. In practice, financial markets are continuously affected by changing investor sentiment, economic policies, international events and unexpected market reactions, making their behaviour highly dynamic and difficult to model using conventional techniques alone.

Forecasting financial time series has therefore remained an important area of research for both academia and industry. Reliable prediction of stock market behaviour can assist investors and financial institutions in making informed decisions related to risk management, investment planning and portfolio optimization. However, the uncertain nature of financial systems often leads to complex temporal patterns that are not adequately captured by traditional linear models. As a result, there has been growing interest in developing more flexible and data driven approaches capable of learning intricate nonlinear relationships directly from observed financial data.

Recent advances in deep learning have significantly improved the capability of predictive models by enabling the extraction of complex nonlinear patterns from sequential data. Neural network architectures have demonstrated promising performance in various forecasting tasks because of their ability to automatically learn hidden structures and temporal dependencies from large datasets. Unlike conventional statistical models, deep learning methods are not strictly constrained by linear assumptions and can therefore adapt more effectively to the complicated behaviour exhibited by financial time series. These developments have motivated researchers to explore deep neural networks for stock market forecasting and related financial applications.

However, many conventional deep learning architectures are deterministic in nature and may fail to adequately represent the stochastic behaviour inherent in financial systems. Financial markets are naturally uncertain environments where randomness plays a fundamental role in the evolution of prices and volatility. Small external disturbances or unforeseen events can lead to substantial fluctuations in market behaviour, making deterministic models insufficient for fully capturing the underlying dynamics of financial systems. Consequently, there is a need for models that can explicitly incorporate uncertainty and randomness within the learning process itself.

Stochastic differential equation based neural networks provide a promising alternative by modeling the hidden dynamics of neural networks through continuous stochastic processes. In particular, SDE-Net incorporates Brownian motion to represent uncertainty in the evolution of hidden states. By introducing stochastic perturbations into the hidden dynamics of neural networks, such models attempt to provide a more realistic representation of noisy and uncertain temporal systems such as financial markets.

Although Brownian motion based models are effective in capturing continuous random fluctuations, they may not sufficiently describe discontinuous jumps frequently observed in stock market behaviour. Financial time series often contain extreme events such as sudden crashes, unexpected price surges and abrupt volatility changes, indicating the presence of non-Gaussian characteristics in the data. Lévy processes, which generalize Brownian motion by allowing jump discontinuities, may therefore provide a more realistic stochastic framework for modeling such phenomena. Their ability to incorporate both continuous fluctuations and sudden jumps makes them potentially more suitable for representing real world financial behaviour.

The primary objective of this project is to investigate whether Lévy driven neural differential equation models can improve forecasting performance over traditional Brownian motion driven SDE-Net models. Using real financial datasets from the NIFTY50 and SSE50 indices, the project aims to compare the predictive capabilities of these stochastic neural architectures under multiple forecasting horizons. The study further seeks to analyze the effectiveness of different stochastic perturbations in capturing the complex and uncertain dynamics of financial markets. Through this comparative analysis, the project attempts to explore the role of stochastic modeling in enhancing modern deep learning based approaches for financial time series forecasting.