\documentclass[11pt, a4paper]{article}

\usepackage[top = 1 in, bottom = 1 in, left = 1 in, right = 1 in]{geometry}

\usepackage{amsmath, amssymb, amsfonts}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tabularray}
\usepackage{undertilde}


\title{readme}
\author{}
\date{}


\begin{document}

\maketitle

Here are some comments/messages that I wanted to convey regarding the notes that are being shared.

\section{MSMS 101}

\begin{itemize}
\item \textit{Part 1 - Variance Bounds} : The last problem is left incomplete. If you have a complete solution or any suggestion, please drop an email to \href{mailto:ami.ananda@bhu.ac.in}{ami.ananda@bhu.ac.in}
	
\item \textit{Part 2 - Sufficiency and Siblings} : Basu's Theorem is not included in the notes.
	
\item \textit{Part 3 - Maximum Likelihood Estimation}
	
\item \textit{Part 4 - Decision Theory}
\end{itemize}


\section{MSMS 102}

\begin{itemize}
\item \textit{Part 1 - Measure Theory}
	
\item \textit{Part 2 - Probability Theory} : The proof of the converse part is left incomplete at page number 8. Sir told us to try the proof on our own. But I think there is an issue with the counter-example or I'm dumb. Anyways I couldn't complete the proof. If you have any suggestion, please drop an email to \href{mailto:ami.ananda@bhu.ac.in}{ami.ananda@bhu.ac.in}
	
\item \textit{Problems} : It has some problems solved from previous year question papers. You can also solve exercises of Rohatgi-Saleh and B.R.Bhatt.
		
\item \textit{Add-ons from DK Sir} : Throughout the semester, DK Sir takes digressions to various concepts/topics of Mathematics and Statistics. This is a non-exhaustive collection of those discussions.
		
\item \textit{Modern Probability Theory - B R Bhat - Problems} : I also solved a few more problems, but they are not well-written.
\end{itemize}


\section{MSMS 103}

\begin{itemize}
\item \textit{Sampling Techniques}

\item \textit{Analysis of Covariance}
	
\item \textit{Design of Experiments} : These are non-exhaustive set of classnotes.
\end{itemize}

\section{MSMS 104}

\begin{itemize}
\item \textit{Numerical Analysis} : Mainly contains derivations of various iterative methods.
\end{itemize}


\end{document}